Analyzing Trends in Crime Against Persons: Insights from the Toronto Police Annual Statistical Report

### 2024-01-24 outputs/paper/paper.qmd
Updated File

library(tidyverse)
library(AER)
library(opendatatoronto)
library(dplyr)
library(readr)
library(janitor)
library(ggplot2)

# Loading Clean data & Read data
cleaned_victim_data <- read_csv(cleaned_data, file="outputs/data/cleaned_lead_data.csv")
```

@@

# Data

# References

>>>>

library(tidyverse)
library(AER)
library(opendatatoronto)
library(dplyr)
library(readr)
library(janitor)
library(ggplot2)
library(knitr)

raw_data <- read_csv(here::here("inputs/data/raw_data.csv"))

cleaned_dataset <-
  read_csv(
    here::here("outputs/data/cleaned_data.csv")
  )
```

@@

# Data {#sec-2}

## Data Source

The data used in this paper was collected by the OpenDataToronto Library [@citeODT]. The specific dataset used in this research is the
'Police-Annual-Statistical-Report-Victims-of-Crimes' [@citeDataset] from the year 2014 to 2022. The dataset is categorized by crime subtype against a person, so basically four-types: Assault, Robbery, Sexual Violation, and Other offenses. The stratification of the dataset follows victim demographics such as age group, age cohort and gender.

## Data Cleaning
```{r}
#| label: tbl-Cleaned-data-output
#| tbl-cap: Sample of Cleaned Victim Dataset 
#| echo: false
head(cleaned_dataset) |>
  kable(
    col.names = c("ID", "YEAR", "CRIME", "SEX", "AGE_GROUP", "AGE_COHORT", "VICTIM_COUNT"),
    booktabs = TRUE
    )
```

Data used in this paper was downloaded, cleaned and analyzed with the programming language R [@citeR]. Also with support of additional packages in R: `tidyverse` [@citeTidyverse], `ggplot2` [@citeGgplot], `janitor` [@citeJanitor], `dplyr` [@citeDplyr], `readr` [@citeReadr], `knitr` [@citeknitr]. The Table 1 showcases the cleaned dataset from the raw data extraced from the 'Police-Annual-Statistical-Report-Victims-of-Crimes' [@citeDataset].

# Results & Analysis

## Total Victims of Crime over Year

Refer to @tbl-Total-Victims-by-Year which shows the data on totals victims of crime over the period of 2014 to 2022 and corresponding to that a line graph is created, both showcase the annual trend in the number of reported victims. @tbl-Total-Victims-by-Year, titled "Total Victims From 2014-2022", is shown in the tabular format, allowing for quick reference to the absolute numbers of victims per year.

```{r}
#| label: tbl-Total-Victims-by-Year
#| tbl-cap: Total Victims From 2014-2022
#| echo: false
summarized_data <- cleaned_dataset %>%
  group_by(report_year) %>%
  summarize(Total_Count = sum(victim_count))

# Use kable from knitr to create a simple table

summarized_data |>
  kable(
    col.names = c("YEAR", "TOTAL VICTIMS OF CRIME"),
    booktabs = TRUE
    )  
```

```{r}
#| label: fig-ttl-victims
#| fig-cap: Total Victims of Crime over Year (2014-2022)
#| echo: false
#| warning: false
#| message: false

# Creating the line graph
ggplot(summarized_data, aes(x = report_year, y = Total_Count)) +
  geom_line(color = "blue") +  # This creates the line graph
  geom_point() +  # Optional: adds points at each year
  labs(title = "Total Victims of Crime",
       x = "Year",
       y = "Total Victims of Crime Over Year") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))  # Rotate x-axis labels for better readability

```

From @fig-ttl-victims - the line graph provides a visual representation of these figures, plotting the total victims of crime over the years. The trend shows a steady increase in Total victim counts from 2014 to 2019, suggesting an increasing rate of victimization during this period and increasing crime rate. However, there is a sharp decline in the year 2020 that invites questions about the potential causes. That was followed by an upward trajectory in 2021 and 2022.

## Demograpic of Victims
